### 2014-02-07 WriteUp/liquidity.tex
related works section; bib files; liquidity write-up

\documentclass[prodmode,acmec]{ec-acmsmall} % Aptara syntax

% Package to generate and customize Algorithm as per ACM style
\usepackage[ruled]{algorithm2e}
\renewcommand{\algorithmcfname}{ALGORITHM}
\SetAlFnt{\small}
\SetAlCapFnt{\small}
\SetAlCapNameFnt{\small}
\SetAlCapHSkip{0pt}
\IncMargin{-\parindent}

% Metadata Information
\acmVolume{9}
\acmNumber{4}
\acmArticle{39}
\acmYear{2013}
\acmMonth{6}
\usepackage[numbers]{natbib}
\usepackage{amsmath}
\usepackage{amssymb}
%\usepackage{amsthm}
\usepackage{parskip}
\usepackage{booktabs}
%new commands
\newcommand{\la}{\langle}
\newcommand{\ra}{\rangle}
\newcommand{\mX}{\mathcal{X}}
\newcommand{\bR}{\mathbf{R}}
\newcommand{\grad}{\nabla}
\newcommand{\Exp}{\mathbf{E}}
\newcommand{\Var}{\mathrm{Var}}
\newcommand{\q}{\mathbf{q}}
\newcommand{\mM}{\ensuremath{\mathcal{M}}}
\newcommand{\hmu}{\ensuremath{\hat{\mu}}}
\newcommand{\mP}{\ensuremath{\mathcal{P}}}
\newcommand{\mY}{\ensuremath{\mathcal{Y}}}
\newcommand{\mS}{\ensuremath{\mathcal{S}}}
\newcommand{\ds}{\displaystyle}
\newcommand{\mF}{\mathcal{F}}
\newcommand{\mD}{\mathcal{D}}
\newcommand{\htheta}{\hat{\theta}}
%---------------------------------------------------------------------------
\newtheorem{thm}{Theorem}[section]
\newtheorem{prop}[thm]{Proposition}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{con}[thm]{Conjecture}
\newtheorem{cor}[thm]{Corollary}
\newtheorem{ex}[thm]{Example}
\newtheorem{qn}[thm]{Question}
\newcommand{\defn}{\bigbreak\noindent{\bf Definition. }}
\newcommand{\undefn}{\bigbreak}
\newcommand{\rmk}{\bigbreak\noindent{\bf Remark. }}
\newcommand{\unrmk}{\bigbreak}
%\newcommand{\proof}{\noindent{\bf Proof. }}
\newcommand{\unproof}{\hfill$\Box$\bigbreak}
\newcommand{\defeq}{\stackrel{\mathrm{def}}{=}}
\newcommand{\holds}[1]{\stackrel{?}{#1}}
\newcommand{\conv}{\mathrm{conv}}
\newcommand{\ie} {{\it i.e. }}
\newcommand{\eg} {{\it e.g. }}
\newcommand{\etal} {{\it et al. }}
\newcommand{\E}{\mathbf{E}}
\newcommand{\betavec}{\pmb{\beta}}
\newcommand{\qvec}{\mathbf{q}}
\newcommand{\lpf}{\mathbf{\psi}} %logpartition function
\newcommand{\family}{\mathcal F} 
\newcommand{\suff}{\pmb{\phi}} % suff stats
\newcommand{\muvec}{\pmb{\mu}}

% Document starts

\begin{document}
\markboth{Abernethy et al.}{Maximum Entropy Prediction Markets}
% Title portion
\title{Maximum Entropy Prediction Markets}
\author{JACOB ABERNETHY
\affil{University of Michigan, Ann Arbor}
SINDHU KUTTY
\affil{University of Michigan, Ann Arbor}
S\'{E}BASTIEN LAHAIE
\affil{Microsoft Research, New York City}
RAHUL SAMI
\affil{University of Michigan, Ann Arbor}}

\section{Cost Function with a Liquidity Parameter}
We may introduce a liquidity parameter to the cost function as $C_{\lambda}(\theta)=\lambda C\left(\frac{\theta}{\lambda}\right)$. We would like to find an expression for a lower bound on expected payoff. We would also like to find an expression for final market state for exponential utility traders with exponential family beliefs.
\subsection{lower bound on expected payoff}
The prediction market is set up with cost function $C_{\lambda}(\theta)$  and payoff for a unit share vector as $\phi(x)$. To find a lower bound on expected payoff, we first note that
\begin{eqnarray*}
\E_{x\sim p(x;\beta)}[C_{\lambda}(\theta)-\theta\phi(x)]&=&C_{\lambda}(\theta)-\theta\nabla T(\beta)\\
&=&\lambda T(\theta/\lambda)-\theta\nabla T(\beta)\\
&=&\lambda [T(\theta/\lambda)-\theta/\lambda\nabla T(\beta)-T(\beta)+\beta\nabla T(\beta)]+\lambda T(\beta)-\theta\lambda\nabla T(\beta)\\
&=&\lambda D_{T}(\theta/\lambda,\beta)+\lambda [T(\beta)-\theta\nabla T(\beta)]
\end{eqnarray*}¥ 
where $D_{T}(x_{1},x_{2}) = T(x_{1})-T(x_{2})-\nabla T(x_{1}-x_{2})$ is the Bregman Divergence between $x_{1}$ and $x_{2}$ based on $T$.

Thus, the expected payoff  for a trader with exponential family belief parameter $\beta$ who moves the market state from $\theta$ to $\theta'$ can be written as
\begin{eqnarray*}
\E_{x\sim p(x;\beta)}\left[C_{\lambda}(\theta)-\theta\phi(x)-(C_{\lambda}(\theta')-\theta'\phi(x))\right]=\lambda[D_{T}(\theta/\lambda,\beta)-D_{T}(\theta'/\lambda,\beta)]
\end{eqnarray*}¥
If $\theta'=a\beta+(1-a)\theta$ for some $a\in[0,1]$ we can write
\begin{eqnarray*}
D_{T}(\theta'/\lambda,\beta)&\leq& a D_{T}(\beta/\lambda,\beta)+(1-a) D_{T}(\theta/\lambda,\beta)\\
D_{T}(\theta/\lambda,\beta)-D(\theta'/\lambda,\beta)&\geq& a D_{T}(\theta/\lambda,\beta)-a D_{T}(\beta/\lambda,\beta)
\end{eqnarray*}¥
%Note that 
%\begin{eqnarray*}
%\nabla_{\theta} C_{\lambda} (\theta)&=&\nabla_{\theta}\lambda C\left(\frac{\theta}{\lambda}\right)\\
%&=&\nabla_{\theta/\lambda}C\left(\frac{\theta}{\lambda}\right)\\
%&=&\E_{\theta/\lambda}[\phi(x)]
%\end{eqnarray*}¥
\subsection{exponential utility traders with exponential family beliefs}
As before the prediction market is set up with cost function $C_{\lambda}(\theta)$  and payoff for a unit share vector as $\phi(x)$. Thus, the  payoff of a trader who moves the market state from $\theta$ to $\theta'$ is given by $C_{\lambda}(\theta)-C_{\lambda}(\theta')+\delta\phi(x)$ where $\delta=\theta'-\theta$ and the expected utility of the payoff for a trader with exponential belief parameter $\beta$ is given by
\begin{eqnarray*}
&&\E_{x\sim p(x;\beta)}U[C_{\lambda}(\theta)-C_{\lambda}(\theta')+\delta\phi(x)]\\
&=&U[C_{\lambda}(\theta)-C_{\lambda}(\theta+\delta)+\frac{1}{a}(T(\beta)-T(\beta-a\delta))]\\
&=&U[\lambda C(\theta/\lambda)-\lambda C\left(\frac{\theta+\delta}{\lambda}\right)+\frac{1}{a}(T(\beta)-T(\beta-a\delta))]\\
&=&U[\lambda T(\theta/\lambda)-\lambda T\left(\frac{\theta+\delta}{\lambda}\right)+\frac{1}{a}(T(\beta)-T(\beta-a\delta))]
\end{eqnarray*}¥
Here $T$ is the log partition function; for the exponential family prediction market $C=T$. To maximize the utility, we set the gradient of the argument with respect to $\delta$ to 0. 
$$\nabla T\left(\frac{\theta+\delta}{\lambda}\right) = \nabla T(\beta-a\delta)$$
So we have for an optimal choice of $\delta$, 
\begin{eqnarray*}
\frac{\theta+\delta}{\lambda} &=& \beta-a\delta\\
\implies \delta&=&\frac{\beta\lambda-\theta}{1+a\lambda}\\
\implies \theta+\delta&=&\lambda\frac{\beta+a\delta}{1+a\lambda}
\end{eqnarray*}

\end{document}