### 2014-04-08 WriteUp/response.txt
edits needed for camera-ready version

- Reduce complications in notation: Please only keep necessary notation in this paper. 
- intro revision: we meant this comment for the narrow case of MSR-style markets.
  - Ostrovsky and of Wolfers and Zitzewitz are exactly about these issues. Pennock, in his Ph.D. thesis (1999), asks the question of what happens in equilibrium when traders have different priors and budget constraints, and shows that in equilibrium, the price can be interpreted as a budget-weighted average of trader beliefs. Beyond this, most of the empirical work on prediction markets is asking exactly the question of what prediction market prices actually mean. The recent working paper of Rothschild and Sethi ("Trading Strategies and Market Microstructure: Evidence from a Prediction Market") discusses how to interpret prices in the context of a continuous market rather than equilibrium prices -- the references in that paper are also worth looking at. Much of the work on manipulation is also implicitly focused on exactly this question (including the paper "Gaming Prediction Markets" with one of the authors of this paper as a co-author!)
- If accepted, the authors should spend some effort in discussing the connection of this work to that of Abernethy-Frongilllo (COLT 2012)
  - The reviewer makes a really good point regarding our omission of a reference to [Abernethy-Frongillo'12, Thm 11]; this work is certainly related although perhaps not to the extent suggested. It is correct that [AF'12, Thm 11] is the right result to use in the 'only if' part of Thm 2.2 (we will definitely include this), which follows up on prior work of [Savage'71] and [Gneiting-Raftery'07]. However, Thm 2.2 is not an instance of or equivalent to any result in [AF'12]; the construction in Thm 2.2 is to apply a scoring rule to a parametrized family of densities. One finds a "family of densities" constructed in the proof of [AF'12, Thm 11], but this does not fall within the exponential family setting.
    - That paper shows how to construct a proper scoring rule for linear properties from any convex function using Bregman divergence. This paper seems to basically present a way of choosing that convex function; this is not uninteresting in itself, but I'm concerned as to why this connection is not made clearer in the paper.
- Highlight key contributions: the authors could have just focused on presenting characterization of the class of new proper scoring rules and behavior and analysis of agents with exponential utility function. The connection between exponential family and the proposed prediction market setting is interesting and helpful if it is briefly mentioned and discussed in the end of the paper. The paper in its current form reads like suggesting that using exponential families is justified at the modeling level, which sounds a little weak (and in fact not necessary). So I dont think they should be mentioned in the beginning. For example, I got confused by "outstanding shares". Also not all exponential family distributions have a mean parameter (which is fine if you choose to model the market price distribution in this way, but this may limit the generality of the framework).
- Please give a formal definition of exponential family distributions (with citation) and examples of commonly studied distributions that can be modeled by the exponential family studied in this paper. 
- The importance of the underlying measure v is understated. If v is always the Lebesgue measure then many distributions cannot be modeled by the exponential family as presented in the paper, e.g. normal distributions with fixed variance. Many discussions and proofs are made without reference to v. My question is: is it true that all the results and comments can be applied to arbitrary v? (I just checked a few but not all since many calculations are already oversimplified.)
  -  Please give an example of non-Lebesgue v.
  - Review 3: The reviewer makes a good point, we did not really address alternative base measure, and we will need to add more detail here. However, all results hold for any arbitrary base measure v. Often v does not appear because we work directly with the expectation mu and do not write out the implicit integral. All integrals are taken with respect to dv, e.g. (5).
- For example, I found the reference to Bergman divergence a little annoying. I also think that maximum entropy principle is overemphasized---proving that the proposed scoring rule is proper is enough to me. I also got lost at the last paragraph of Section 1.
- P3: Section 2, after first paragraph. Please clarify that P contains ALL distributions on the outcome space.
- P4: please clarify the setup. Are u always using a single outcome x?, same in the example section at Page 7
- I found the "underlying density" and "full density" discussion on P4 very confusing (what are these two densities?).
- P4, last sentence. Please quote what exactly was proven by [Grunwald and Dawid]
- P5: sufficient statistics - please clarify which distributions are not like this. Can it be that the exponential distribution is not like this?
  - Review 1: The concept of a sufficient statistic also arises for distributions other than exponential families. An example of a non-exponential family with sufficient statistic is the uniform distribution over [0,r], where r is the parameter. The sufficient statistic for r is the maximum of the observations.
- P6, THM 2.2. Where the sufficient statistics condition fits in. Can it be that the entropy maximizing exponential distribution is not like that. Please clarify.
- P7: U are using mu for both the report and expectation. Please use a different notation (e.g. hat(mu)
- Can you model classical prediction markets with finite outcome space under the proposed framework? If so then what distribution in the exponential family its correspond to?
  - Classical prediction markets (specifically, LMSR) can be recovered from our framework by using the multinomial distribution, which is an exponential family. It was originally covered in the examples of Sections 2.2 and 3.3 but removed due to space constraints.
- Section 5: Please clarify the setup up-front. What is known to the principal on the traders? Does it know $a$ or it can be estimated
  - In Section 5, the principal (market maker) does not need to know the risk aversion parameter 'a' for Thms 5.2 and 5.3 to hold. The aggregation arises directly from the agents' choice of trades.
- What kind of equilibria are studied in Section 5.2? I.e. what is the game (especially what are the strategies)? What is the dynamics under the game? How equilibria are reached from the initial state of the market?
  - we should add a few more details regarding the results of 5.2, but here are some short answers: The utility functions are defined in 5.2, the strategies are the vectors of share purchases of each agent, we are interested in pure-strategy Nash equilibria, and the dynamics are not required to establish the existence of an equilibrium. We will expand upon these in the revision.
- Section 6: What exactly does the statement "the traders trade in multiple instances of the market" mean? What are the formal definitions of an "uninformative" and an "informative" trader? The semantics of the time-varying budget alpha_i^t are not clear.
- Regarding budgets and play money: indeed the play-money markets are a case where budget constraints are key. We have experience with this in other work where a play-money market was released for the US Elections. Here the budget constraints help alleviate the problem that traders can easily manipulate prices with unreasonable trades -- the goal of Section 6 was to give formal results to this effect. Risk aversion is also plausible with play-money; for instance, a trader high on the leaderboard may not want to risk her standing with large late trades.
  - the example for risk-aversion of someone not wanting to lose their place on the leaderboard doesn't really hold water: that's a type of tournament incentive: very different from traditional risk-aversion.